Effects of Fishing on Scarid Grazing Behavior in the Caribbean

### 2018-03-15 DonorReport.Rmd
updates, adding new summary file for orals

```{r,message=FALSE,warning=FALSE, echo=FALSE}
# set up
library(rmarkdown)
library(dplyr)
library(tidyr)
library(lazyeval)
library(ggplot2)
library(ellipse)
library(RColorBrewer)
library(ggbiplot)
library(nlme)
library(pls)
library(mgcv)
library(ggsignif)
```
```{r,message=FALSE,warning=FALSE, echo=FALSE}
setwd("~/github/Scarid-Behavior/data")
site.sum <- read.csv("site.sum.csv", stringsAsFactors=FALSE) %>% select(-X)
data <- read.csv("sumID2.csv", stringsAsFactors=FALSE) %>% select(Island,Site,Spearfishing,Diving,Species,Species.Code,Phase,Length.cm,g.frac,FR,BR,TDB,for.dur,for.bites,tot.BM,tot.DEN,scar.BM,herb.BM,scar.DEN,conspec.BM,conspec.DEN,carn.BM,carn.DEN,cover_CCA,cover_LC,cover_MA,cover_TA,canopy_MA,canopy_TA,rugosity,int.tot,int.con,int.hetscar,int.dam)
vetvir <- data %>% filter(Species.Code!="rbp")
vetvir20 <- vetvir %>% filter(Length.cm>=15 & Length.cm<=25)
init20<- vetvir %>% filter(Length.cm>=15 & Length.cm<=25 & Phase=="i")
vetvirAB <- vetvir %>% filter(Island!="Barbuda")
vetvirAB20 <- vetvir %>% filter(Island!="Barbuda")
vetvirAB20NS <- vetvirAB20 %>% filter(Site!="Snapper Point")
vetvir20NS <- vetvir20 %>% filter(Site!="Snapper Point")
vetvirNS <- vetvir %>% filter(Site!="Snapper Point")
```

@@

```{r, message=FALSE,warning=FALSE,echo=FALSE}
ggplot(vetvir20, aes(x=Island, y=FR)) + geom_boxplot() + facet_grid(.~Species) +
stat_summary(fun.y = mean, geom="point", shape=5, size=3) + labs(y="Feeding Rate (bites/hr)",x="",title="Feeding Rate by Island and Species") + theme_bw() + theme(plot.title = element_text(hjust = 0.5))
```
```{r, message=FALSE,warning=FALSE,echo=FALSE}
# do I really satisfy ANOVA assumptions?
vet<- vetvir %>% filter(Species.Code=="qup")
vir<- vetvir %>% filter(Species.Code=="stop")
aovFRvet <- aov(FR ~ Island, data=vet) 
#TukeyHSD(aovFRvet)
aovFRvir <- aov(FR ~ Island, data=vir) 
#TukeyHSD(aovFRvir)
```

@@

```{r, message=FALSE,warning=FALSE,echo=FALSE}
ggplot(vetvir20, aes(x=Island, y=for.bites)) + geom_boxplot()  + labs(y="Bites per foray",x="",title="Grazing intensity by Island and Species") + theme_bw() + theme(plot.title = element_text(hjust = 0.5)) + geom_signif(comparisons = list(c("Antigua", "Barbuda"),c("Antigua","Bonaire"),c("Barbuda","Bonaire")), map_signif_level=TRUE) + facet_grid(.~Species)
```
```{r, message=FALSE,warning=FALSE,echo=FALSE}
# do I really satisfy ANOVA assumptions?
aovforvet <- aov(for.dur ~ Island, data=vet) 
#TukeyHSD(aovforvet)
aovforvir <- aov(for.dur ~ Island, data=vir) 
#TukeyHSD(aovforvir)
```
Grazing intensity, or the number of consecutive bites taken during a feeding foray, has significant implications for the benthic substrate. When parrotfish graze more intensely, they are more likely to crop algae and clear space for recruiting or competing corals. When shorter grazing forays are taken, bites are less concentrated and algae may not be sufficiently reduced in any given area. This plot shows the significant difference in grazing intensity of *Scarus vetula* individuals in Bonaire and of those in Antigua and Barbuda (p<0.001). For *Sparisoma viride*, grazing intensity in both Bonaire and Barbuda were significantly greater than in Antigua (p<0.001 and p<0.05, respectively).

>>>>

```{r,message=FALSE,warning=FALSE, echo=FALSE}
# set up
library(rmarkdown)
library(dplyr)
library(tidyr)
library(lazyeval)
library(ggplot2)
library(ellipse)
library(RColorBrewer)
library(ggbiplot)
library(nlme)
library(pls)
library(mgcv)
library(ggsignif)
library(car)
```
```{r,message=FALSE,warning=FALSE, echo=FALSE}
setwd("~/github/Scarid-Behavior/data")
site.sum <- read.csv("site.sum.csv", stringsAsFactors=FALSE) %>% select(-X)
data <- read.csv("sumID2.csv", stringsAsFactors=FALSE) %>% select(Island,Site,Spearfishing,Diving,Species,Species.Code,Phase,Length.cm,g.frac,FR,BR,TDB,for.dur,for.bites,tot.BM,tot.DEN,scar.BM,herb.BM,scar.DEN,conspec.BM,conspec.DEN,carn.BM,carn.DEN,cover_CCA,cover_LC,cover_MA,cover_TA,canopy_MA,canopy_TA,rugosity,int.tot,int.con,int.hetscar,int.dam)
vetvir <- data %>% filter(Species.Code!="rbp")
vetvir20 <- vetvir %>% filter(Length.cm>=15 & Length.cm<=25)
init20<- vetvir %>% filter(Length.cm>=15 & Length.cm<=25 & Phase=="i")
vetvirAB <- vetvir %>% filter(Island!="Barbuda")
vetvirAB20 <- vetvir %>% filter(Island!="Barbuda")
vetvirAB20NS <- vetvirAB20 %>% filter(Site!="Snapper Point")
vetvir20NS <- vetvir20 %>% filter(Site!="Snapper Point")
vetvirNS <- vetvir %>% filter(Site!="Snapper Point")
```

@@

```{r, message=FALSE,warning=FALSE,echo=FALSE}
ggplot(vetvir20, aes(x=Island, y=FR)) + geom_boxplot() + facet_grid(.~Species) +
stat_summary(fun.y = mean, geom="point", shape=5, size=3) + labs(y="Feeding Rate (bites/hr)",x="",title="Feeding Rate by Island and Species") + theme_bw() + theme(plot.title = element_text(hjust = 0.5))
```
```{r, message=FALSE,warning=FALSE,echo=FALSE}
# do I really satisfy ANOVA assumptions?
vet<- vetvir %>% filter(Species.Code=="qup")
aovFRvet <- aov(FR ~ Island, data=vet) 
TukeyHSD(aovFRvet)
plot(aovFRvet, 2) # check normality
leveneTest(FR ~ Island, data = vet) # check homogeneity of variance

vir<- vetvir %>% filter(Species.Code=="stop")
aovFRvir <- aov(FR ~ Island, data=vir) 
TukeyHSD(aovFRvir)
plot(aovFRvir, 2) # check normality
leveneTest(FR ~ Island, data = vir) # check homogeneity of variance
```

@@

```{r, message=FALSE,warning=FALSE,echo=FALSE}
ggplot(vetvir20, aes(x=Island, y=for.bites)) + geom_boxplot()  + labs(y="Bites per foray",x="",title="Grazing intensity by Island and Species") + theme_bw() + theme(plot.title = element_text(hjust = 0.5)) + geom_signif(comparisons = list(c("Antigua", "Barbuda"),c("Antigua","Bonaire"),c("Barbuda","Bonaire")), map_signif_level=TRUE) + facet_grid(.~Species)
```
```{r, message=FALSE,warning=FALSE,echo=FALSE}
# do I really satisfy ANOVA assumptions?
aovforvet <- aov(for.dur ~ Island, data=vet) 
TukeyHSD(aovforvet)
aovforvir <- aov(for.dur ~ Island, data=vir) 
TukeyHSD(aovforvir)
```
Grazing intensity, or the number of consecutive bites taken during a feeding foray, has significant implications for the benthic substrate. When parrotfish graze more intensely, they are more likely to crop algae and clear space for recruiting or competing corals. When shorter grazing forays are taken, bites are less concentrated and algae may not be sufficiently reduced in any given area. This plot shows the significant difference in grazing intensity of *Scarus vetula* individuals in Bonaire and of those in Antigua and Barbuda (p<0.001). For *Sparisoma viride*, grazing intensity in both Bonaire and Barbuda were significantly greater than in Antigua (p<0.001 and p<0.05, respectively).
